---
title: "bistro validation figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(ggupset)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

# set plotting theme
theme_set(theme_bw() +
  theme(
    text = element_text(size = 15),
    strip.background = element_rect(fill = "white", color = "white")
  ))

# matches
provedit_exact_matches <-
  read_csv("output/provedit/exact_matches.csv") %>%
  separate(human_id,
    sep = "-",
    into = c("proj1", "proj2", "id")
  ) %>%
  mutate(
    id = ifelse(nchar(id) == 1, paste0("0", id), id),
    human_id = ifelse(is.na(proj1), NA, paste0(proj1, "-", proj2, "-", id))
  ) %>%
  select(-c(proj1, proj2, id))

provedit_similarity_matches <-
  read_csv("output/provedit/similarity_matches.csv") %>%
  separate(human_id,
    sep = "-",
    into = c("proj1", "proj2", "id")
  ) %>%
  mutate(
    id = ifelse(nchar(id) == 1, paste0("0", id), id),
    human_id = ifelse(is.na(proj1), NA, paste0(proj1, "-", proj2, "-", id))
  ) %>%
  select(-c(proj1, proj2, id))

provedit_similarity <-
  read_csv("output/provedit/human_similarities.csv")
provedit_n_hu <-
  n_distinct(c(provedit_similarity$hu1, provedit_similarity$hu2))
provedit_max_similarity <- provedit_similarity %>%
  pull(similarity) %>%
  max()

provedit_bistro_matches <-
  read_csv("output/provedit/bistro_matches.csv") %>%
  mutate(
    match = str_to_sentence(match),
    incomplete = case_when(
      locus_count == 15 ~ "Complete profile",
      locus_count < 15 ~ "Incomplete profile"
    )
  )

provedit_bistro_matches_locus <-
  read_csv("output/provedit/bistro_matches_locus.csv") %>%
  mutate(match = str_to_sentence(match))

provedit_thresh10_matches <-
  read_csv("output/provedit/bistro_log10LRs.csv") %>%
  mutate(
    match = ifelse(log10_lr >= 10 &
      !is.infinite(log10_lr), "Yes", "No"),
    human_id = ifelse(match == "Yes", human_id, NA),
    log10_lr = ifelse(match == "Yes", log10_lr, NA)
  ) %>%
  select(bloodmeal_id, match, human_id, log10_lr) %>%
  distinct()

bm_similarity <-
  read_csv("output/bloodmeals/human_similarities.csv")
bm_n_hu <- n_distinct(c(bm_similarity$hu1, bm_similarity$hu2))
bm_max_similarity <- bm_similarity %>%
  filter(similarity != 1) %>%
  pull(similarity) %>%
  max()

bm_exact_matches <- read_csv("output/bloodmeals/exact_matches.csv")
bm_similarity_matches <-
  read_csv("output/bloodmeals/similarity_matches.csv")

bm_bistro_matches_all <-
  read_csv("output/bloodmeals/bistro_matches.csv")

bm_bistro_matches <- bm_bistro_matches_all %>%
  filter(notes != "all log10LRs NA or Inf;no peaks above threshold") %>%
  mutate(
    match = str_to_sentence(match),
    incomplete = case_when(
      locus_count == 9 ~ "Complete profile",
      locus_count < 9 ~ "Incomplete profile"
    ),
    multisource = case_when(
      est_noc == 1 ~ "Single-source",
      est_noc > 1 ~ "Multi-source"
    )
  )

bm_lrs <- read_csv("output/bloodmeals/bistro_log10LRs.csv")
bm_info <- bm_lrs %>%
  select(bloodmeal_id, locus_count, est_noc) %>%
  distinct()

bm_thresh10_matches_all <- bm_lrs %>%
  filter(log10_lr >= 10 & !is.infinite(log10_lr)) %>%
  mutate(match = "Yes") %>%
  select(bloodmeal_id, match, human_id)

bm_thresh10_matches_all <- bm_info %>%
  left_join(bm_thresh10_matches_all) %>%
  mutate(match = ifelse(is.na(match), "No", match))

bm_thresh10_matches <- bm_thresh10_matches_all %>%
  filter(bloodmeal_id %in% bm_bistro_matches$bloodmeal_id)
```

```{r}
provedit_samp_composition <- provedit_exact_matches %>%
  select(bloodmeal_id) %>%
  rowwise() %>%
  mutate(
    props_orig = gsub(
      "^[0-9]+-", "",
      gsub(".*[0-9][0-9]_|-M.*", "", bloodmeal_id)
    ),
    props_new = gsub("-", ";", props_orig),
    first_part = gsub(paste0(props_orig, "-M.*"), "", bloodmeal_id),
    second_part = gsub(paste0(".*", props_orig), "", bloodmeal_id),
    bloodmeal_id_alt = ifelse(
      grepl("M", bloodmeal_id),
      paste0(first_part, props_new, second_part),
      bloodmeal_id
    )
  ) %>%
  ungroup() %>%
  select(-c(props_orig, props_new, first_part, second_part)) %>%
  separate(
    bloodmeal_id_alt,
    sep = "-",
    into = c(
      "project1",
      "project2",
      "contributor_id_dilution",
      "amplification_or_mixture",
      "experiment_or_mixture",
      "amplification_mixture",
      "experiment_mixture"
    ),
    remove = FALSE
  ) %>%
  separate(
    contributor_id_dilution,
    sep = "d",
    into = c("human_id", "dilution_condition")
  ) %>%
  mutate(
    project = paste0(project1, "-", project2),
    dilution_number = ifelse(
      is.na(dilution_condition),
      gsub("[a-z]|[A-Z]", "", experiment_or_mixture),
      gsub("[a-z]", "", dilution_condition)
    ),
    condition = ifelse(
      is.na(dilution_condition),
      gsub("M[0-9]", "", experiment_or_mixture),
      gsub("[0-9]", "", dilution_condition)
    ),
    experiment = ifelse(
      is.na(dilution_condition),
      experiment_mixture,
      experiment_or_mixture
    ),
    template_mass = ifelse(
      is.na(dilution_condition),
      amplification_mixture,
      amplification_or_mixture
    ),
    mixture_ratio = ifelse(
      amplification_or_mixture == template_mass,
      NA,
      amplification_or_mixture
    ),
    template_mass = as.numeric(gsub("IP", "", template_mass))
  ) %>%
  separate_rows(human_id, mixture_ratio) %>%
  mutate(
    human_id = gsub(".*_", "", paste0(project, "-", human_id)),
    mixture_ratio = as.numeric(mixture_ratio)
  ) %>%
  group_by(bloodmeal_id) %>%
  mutate(
    noc = n(),
    mixture_proportion = ifelse(is.na(mixture_ratio), 1, mixture_ratio / sum(mixture_ratio)),
    multisource = ifelse(noc == 1, "Single-source", "Multi-source")
  ) %>%
  select(
    -c(
      # duplicates
      project1,
      project2,
      dilution_condition,
      experiment_or_mixture,
      experiment_mixture,
      amplification_or_mixture,
      amplification_mixture,
      project,
      # unneeded
      dilution_number,
      condition,
      experiment,
      mixture_ratio
    )
  ) %>%
  ungroup() %>%
  left_join(
    provedit_bistro_matches %>%
      select(bloodmeal_id, locus_count, incomplete) %>%
      distinct()
  )
```

## Figure 1: Matching algorithm

```{r}
bistro_algorithm <- grViz(
  "digraph bistro {

  # nodes
  node [shape = box,
  fontname = Helvetica]
  'log10_lrs for all\nbloodmeal-human pairs';
  'Select bloodmeal';
  'All log10_lrs < 1.5?';
  'Set threshold\nt = floor(max(log10_lr)*2)/2';
  'Decrement threshold by 0.5';
  'Count log10_lrs \u2265 t';
  'Does output match t + 0.5?';
  't = 1?';
  'No matches';
  'Final matches for bloodmeal';
  'More bloodmeals?';
  'End'

  # edge statements
  edge [fontname = Helvetica]
  'log10_lrs for all\nbloodmeal-human pairs' -> 'Select bloodmeal'
  'Select bloodmeal' -> 'All log10_lrs < 1.5?'
  'All log10_lrs < 1.5?' -> 'No matches' [label = ' Yes']
  'All log10_lrs < 1.5?' -> 'Set threshold\nt = floor(max(log10_lr)*2)/2' [label = ' No']
  'Set threshold\nt = floor(max(log10_lr)*2)/2' -> 'Count log10_lrs \u2265 t'
  'Decrement threshold by 0.5' -> 'Count log10_lrs \u2265 t'
  'Count log10_lrs \u2265 t' -> 't = 1?' [label = '< NOC']
  'Count log10_lrs \u2265 t' -> 'Does output match t + 0.5?' [label = '≥ NOC']
  'Does output match t + 0.5?' -> 't = 1?' [label = 'No']
  'Does output match t + 0.5?' -> 'Select largest\nmatch set pair ≤ NOC' [label = 'Yes']
  't = 1?' -> 'Decrement threshold by 0.5' [label = 'No']
  't = 1?' -> 'Select largest\nmatch set pair ≤ NOC' [label = 'Yes']
  'No matches' -> 'More bloodmeals?'
  'Select largest\nmatch set pair ≤ NOC' -> 'No matches' [label = '= 0']
  'Select largest\nmatch set pair ≤ NOC' -> 'Final matches for bloodmeal' [label = '> 0']
  'Final matches for bloodmeal' -> 'More bloodmeals?'
  'More bloodmeals?' -> 'Select bloodmeal' [label = ' Yes']
  'More bloodmeals?' -> 'End' [label = ' No']

  subgraph {
    rank = same; 'Final matches for bloodmeal'; 'No matches'
  }
}
"
)

bistro_algorithm

bistro_algorithm %>%
  export_svg() %>%
  charToRaw() %>%
  rsvg_png("figures/bistro_algorithm.png")
```


## Figure 2: Evaluation with provedit data

Number of human sources

```{r}
provedit_n_hu
```


Number of single- and multi-source samples

```{r}
provedit_samp_composition %>%
  select(bloodmeal_id, multisource) %>%
  distinct() %>%
  group_by(multisource) %>%
  tally()
```

Maximum provedit similarity value between people in the database

```{r}
provedit_max_similarity
```


```{r}
provedit_exact_matches_correct_none <- provedit_samp_composition %>%
  left_join(provedit_exact_matches) %>%
  filter(!is.na(human_id)) %>%
  mutate(match = ifelse(is.na(match), "No", match))

provedit_exact_matches_incorrect <-
  anti_join(provedit_exact_matches, provedit_samp_composition) %>%
  filter(!is.na(human_id) & match == "Yes")

provedit_similarity_matches_correct_none <-
  provedit_samp_composition %>%
  left_join(provedit_similarity_matches) %>%
  filter(!is.na(human_id)) %>%
  mutate(match = ifelse(is.na(match), "No", match))

provedit_similarity_matches_incorrect <-
  anti_join(provedit_similarity_matches, provedit_samp_composition) %>%
  filter(!is.na(human_id) & match == "Yes")

provedit_bistro_matches_correct_none <-
  provedit_samp_composition %>%
  left_join(
    provedit_bistro_matches %>%
      rowwise() %>%
      mutate(human_id = ifelse(
        nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id
      )) %>%
      ungroup() %>%
      select(-c(locus_count, incomplete))
  ) %>%
  filter(!is.na(human_id)) %>%
  mutate(match = ifelse(is.na(match), "No", match))

provedit_bistro_matches_incorrect <-
  anti_join(
    provedit_bistro_matches %>%
      select(-c(locus_count, incomplete)) %>%
      rowwise() %>%
      mutate(human_id = ifelse(
        nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id
      )) %>%
      filter(match == "Yes"),
    provedit_samp_composition
  ) %>%
  filter(!is.na(human_id))

provedit_thresh10_matches_correct_none <-
  provedit_samp_composition %>%
  left_join(provedit_thresh10_matches %>%
    rowwise() %>%
    mutate(human_id = ifelse(
      nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id
    )) %>%
    ungroup()) %>%
  filter(!is.na(human_id)) %>%
  mutate(match = ifelse(is.na(match), "No", match))

provedit_thresh10_matches_incorrect <-
  anti_join(
    provedit_thresh10_matches %>%
      rowwise() %>%
      mutate(human_id = ifelse(
        nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id
      )) %>%
      filter(match == "Yes"),
    provedit_samp_composition
  ) %>%
  filter(!is.na(human_id))
```


```{r}
provedit_bistro_matches_by_samp <-
  provedit_bistro_matches_correct_none %>%
  select(bloodmeal_id, noc, locus_count, match) %>%
  group_by(bloodmeal_id, noc, locus_count) %>%
  mutate(n_match = sum(match == "Yes")) %>%
  select(-match) %>%
  distinct() %>%
  left_join(
    provedit_bistro_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())
  ) %>%
  mutate(n_incorrect = ifelse(is.na(n_incorrect), 0, n_incorrect)) %>%
  mutate(n_missing = noc - n_match - n_incorrect) %>%
  pivot_longer(c(n_match, n_incorrect, n_missing), names_to = "match_type") %>%
  mutate(
    match_type = case_when(
      match_type == "n_match" ~ "Correct match",
      match_type == "n_incorrect" ~ "Incorrect match",
      match_type == "n_missing" ~ "No match"
    )
  ) %>%
  mutate(
    incomplete = case_when(
      locus_count == 15 ~ "Complete\nprofile",
      TRUE ~ "Incomplete\nprofile"
    ),
    multisource = factor(
      ifelse(noc == 1, "Single-source", "Multi-source"),
      levels = c("Single-source", "Multi-source")
    ),
    match_type = factor(match_type, levels = rev(
      c("Correct match", "Incorrect match", "No match")
    ))
  )

(
  provedit_match_plot <- provedit_bistro_matches_by_samp %>%
    filter(match_type == "Correct match") %>%
    ggplot(aes(
      fill = as.character(noc), x = as.character(value)
    )) +
    geom_bar() +
    scale_fill_grey(start = 0.9, end = 0.1) +
    labs(fill = "Number of\ncontributors", y = "Number of\nsamples", x = "Number of correct matches\nidentified by bistro")
)
```


```{r}
provedit_bistro_matches_by_samp %>%
  group_by(multisource) %>%
  mutate(tot = n_distinct(bloodmeal_id)) %>%
  group_by(multisource, tot, match_type) %>%
  summarize(n_match = sum(value > 0)) %>%
  filter(match_type == "Correct match") %>%
  mutate(prop = n_match / tot)

provedit_bistro_matches_by_samp %>%
  group_by(multisource) %>%
  mutate(tot = sum(value)) %>%
  group_by(multisource, tot, match_type) %>%
  summarize(n_match = sum(value)) %>%
  filter(match_type == "Correct match") %>%
  mutate(prop = n_match / tot)
```


```{r}
(provedit_mixes <- provedit_bistro_matches_correct_none %>%
  mutate(match = factor(match, levels = c("Yes", "No"))) %>%
  filter(multisource == "Multi-source") %>%
  ggplot(aes(x = match, y = mixture_proportion)) +
  stat_compare_means(size = 3, label.x.npc = 0.4) +
  geom_boxplot() +
  labs(x = "Match identified\nby bistro", y = "Mixture proportion"))
```

```{r}
provedit_bistro_matches_correct_none %>%
  group_by(multisource, incomplete) %>%
  mutate(tot = n()) %>%
  group_by(match, incomplete, multisource, tot) %>%
  tally() %>%
  filter(match == "Yes") %>%
  mutate(prop = n / tot)

provedit_bistro_matches_correct_none %>%
  group_by(incomplete) %>%
  mutate(tot = n()) %>%
  group_by(match, incomplete, tot) %>%
  tally() %>%
  filter(match == "Yes") %>%
  mutate(prop = n / tot)
```

```{r}
provedit_bistro_matches_incorrect_locus <- anti_join(bind_rows(
  provedit_bistro_matches %>%
    filter(incomplete == "Complete profile") %>%
    select(-c(incomplete)) %>% rename(locus_count_test = locus_count),
  provedit_bistro_matches_locus %>% rename(locus_count_test = locus_count)
) %>%
  rowwise() %>%
  mutate(human_id = ifelse(nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id)) %>%
  filter(match == "Yes" & bloodmeal_id %in% provedit_samp_composition$bloodmeal_id), provedit_samp_composition) %>%
  filter(!is.na(human_id))

provedit_bistro_matches_correct_none_locus <- provedit_samp_composition %>%
  filter(locus_count == 15) %>%
  full_join(bind_rows(
    provedit_bistro_matches %>%
      filter(incomplete == "Complete profile"),
    provedit_bistro_matches_locus
  ) %>%
    rowwise() %>%
    mutate(human_id = ifelse(nchar(human_id) == 11, paste0(substr(human_id, 1, 10), 0, substr(human_id, 11, 11)), human_id)) %>%
    ungroup() %>%
    select(-c(incomplete)) %>%
    rename(locus_count_test = locus_count)) %>%
  mutate(match = ifelse(is.na(match), "No", match)) %>%
  filter(!(is.na(human_id) & match == "Yes")) %>%
  filter(bloodmeal_id %in% provedit_samp_composition$bloodmeal_id) %>%
  group_by(bloodmeal_id) %>%
  mutate(
    noc = unique(noc[!is.na(noc)]),
    multisource = unique(multisource[!is.na(multisource)])
  ) %>%
  anti_join(provedit_bistro_matches_incorrect_locus)

provedit_locus_compare <- provedit_bistro_matches_correct_none_locus %>%
  group_by(bloodmeal_id, noc, locus_count_test) %>%
  summarize(n_match = sum(match == "Yes")) %>%
  ungroup() %>%
  left_join(provedit_bistro_matches_incorrect_locus %>%
    group_by(bloodmeal_id, locus_count_test) %>%
    summarize(n_incorrect = n())) %>%
  mutate(n_incorrect = ifelse(is.na(n_incorrect), 0, n_incorrect)) %>%
  rowwise() %>%
  mutate(n_missing = max((noc - n_match - n_incorrect), 0)) %>%
  ungroup() %>%
  pivot_longer(c(n_match, n_incorrect, n_missing), names_to = "match_type") %>%
  mutate(
    match_type = case_when(
      match_type == "n_match" ~ "Correct match",
      match_type == "n_incorrect" ~ "Incorrect match",
      match_type == "n_missing" ~ "No match"
    ),
    match_type = factor(match_type, levels = rev(c("Correct match", "Incorrect match", "No match")))
  ) %>%
  left_join(provedit_bistro_matches_locus %>% rename(locus_count_test = locus_count) %>%
    select(bloodmeal_id, locus_count_test) %>%
    distinct()) %>%
  mutate(multisource = factor(ifelse(noc == 1, "Single-source", "Multi-source"),
    levels = c("Single-source", "Multi-source")
  )) %>%
  group_by(locus_count_test, match_type) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  add_row(locus_count_test = 1:15, match_type = "True negative") %>%
  group_by(locus_count_test) %>%
  mutate(value = ifelse(match_type == "True negative", sum(provedit_n_hu - provedit_samp_composition$noc) - value[match_type == "Incorrect match"], value)) %>%
  mutate(match_type = case_when(
    match_type == "No match" ~ "False negative",
    match_type == "Incorrect match" ~ "False positive",
    match_type == "Correct match" ~ "True positive",
    TRUE ~ match_type
  )) %>%
  group_by(locus_count_test) %>%
  summarize(
    sensitivity = value[match_type == "True positive"] / (value[match_type == "True positive"] + value[match_type == "False negative"]),
    specificity = value[match_type == "True negative"] / (value[match_type == "True negative"] + value[match_type == "False positive"])
  )
provedit_locus_compare
```

```{r}
(incomplete_sens_spec <- provedit_locus_compare |>
  pivot_longer(c(sensitivity, specificity)) |>
  mutate(name = str_to_sentence(name)) |>
  ggplot(aes(x = locus_count_test, y = value, linetype = name)) +
  geom_point() +
  geom_line() +
  labs(x = "Number of loci used for matching", y = "Value") +
  theme(
    legend.position = c(0.78, 0.15), legend.box.background = element_rect(color = "black"),
    legend.title = element_blank(), legend.margin = margin(c(-4, 5, 5, 5))
  ))
```

```{r}
provedit_comparison <- bind_rows(
  provedit_exact_matches_correct_none %>%
    select(bloodmeal_id, noc, match) %>%
    group_by(bloodmeal_id, noc) %>%
    mutate(n_match = sum(match == "Yes")) %>%
    select(-match) %>%
    distinct() %>%
    left_join(provedit_exact_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Exact"),
  provedit_similarity_matches_correct_none %>%
    select(bloodmeal_id, noc, match) %>%
    group_by(bloodmeal_id, noc) %>%
    mutate(n_match = sum(match == "Yes")) %>%
    select(-match) %>%
    distinct() %>%
    left_join(provedit_similarity_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Similarity"),
  provedit_thresh10_matches_correct_none %>%
    select(bloodmeal_id, noc, match) %>%
    group_by(bloodmeal_id, noc) %>%
    mutate(n_match = sum(match == "Yes")) %>%
    select(-match) %>%
    distinct() %>%
    left_join(provedit_thresh10_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Static threshold"),
  provedit_bistro_matches_correct_none %>%
    select(bloodmeal_id, noc, match) %>%
    group_by(bloodmeal_id, noc) %>%
    mutate(n_match = sum(match == "Yes")) %>%
    select(-match) %>%
    distinct() %>%
    ungroup() %>%
    left_join(provedit_bistro_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "bistro")
) %>%
  mutate(n_incorrect = ifelse(is.na(n_incorrect), 0, n_incorrect)) %>%
  rowwise() %>%
  mutate(n_missing = max((noc - n_match - n_incorrect), 0)) %>%
  ungroup() %>%
  pivot_longer(c(n_match, n_incorrect, n_missing), names_to = "match_type") %>%
  mutate(
    match_type = case_when(
      match_type == "n_match" ~ "Correct match",
      match_type == "n_incorrect" ~ "Incorrect match",
      match_type == "n_missing" ~ "No match"
    ),
    match_type = factor(match_type, levels = rev(c("Correct match", "Incorrect match", "No match")))
  ) %>%
  left_join(provedit_bistro_matches %>%
    select(bloodmeal_id, locus_count) %>%
    distinct()) %>%
  mutate(
    incomplete = case_when(
      locus_count == 15 ~ "Complete\nprofile",
      TRUE ~ "Incomplete\nprofile"
    ),
    multisource = factor(ifelse(noc == 1, "Single-source", "Multi-source"),
      levels = c("Single-source", "Multi-source")
    ),
    algorithm = factor(algorithm, levels = c("Exact", "Similarity", "Static threshold", "bistro"))
  ) %>%
  group_by(algorithm, multisource, incomplete, match_type) %>%
  summarize(value = sum(value))
```

```{r}
provedit_comparison %>%
  group_by(algorithm, match_type) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  add_row(algorithm = "Exact", match_type = "True negative") %>%
  add_row(algorithm = "Similarity", match_type = "True negative") %>%
  add_row(algorithm = "Static threshold", match_type = "True negative") %>%
  add_row(algorithm = "bistro", match_type = "True negative") %>%
  group_by(algorithm) %>%
  mutate(value = ifelse(match_type == "True negative", sum(provedit_n_hu - provedit_samp_composition$noc) - value[match_type == "Incorrect match"], value)) %>%
  mutate(match_type = case_when(
    match_type == "No match" ~ "False negative",
    match_type == "Incorrect match" ~ "False positive",
    match_type == "Correct match" ~ "True positive",
    TRUE ~ match_type
  )) %>%
  group_by(algorithm) %>%
  summarize(
    sensitivity = value[match_type == "True positive"] / (value[match_type == "True positive"] + value[match_type == "False negative"]),
    specificity = value[match_type == "True negative"] / (value[match_type == "True negative"] + value[match_type == "False positive"])
  )
```

```{r}
full_join(
  provedit_comparison %>%
    group_by(algorithm, match_type) %>%
    summarize(value = sum(value)) %>%
    filter(match_type == "Correct match" & algorithm != "bistro"),
  provedit_comparison %>%
    group_by(algorithm, match_type) %>%
    summarize(value = sum(value)) %>%
    filter(match_type == "Correct match" & algorithm == "bistro") %>%
    rename(bistro = value) %>%
    ungroup() %>%
    select(-algorithm)
) %>%
  mutate(
    diff = bistro - value,
    times_increase = 1 + diff / value
  )
```

```{r}
(provedit_compare_matches <- provedit_comparison %>%
  filter(value != 0) %>%
  mutate(
    match_type = gsub(" match", "", match_type),
    match_type = factor(ifelse(match_type == "No", "None", match_type),
      levels = c("None", "Incorrect", "Correct")
    ),
    multisource = factor(multisource, levels = c("Multi-source", "Single-source")),
    algorithm = factor(algorithm, levels = rev(levels(algorithm)))
  ) %>%
  filter(match_type == "Correct") %>%
  group_by(algorithm, multisource) %>%
  summarize(value = sum(value)) %>%
  ggplot(aes(x = algorithm, y = value, fill = multisource)) +
  geom_col(width = 0.5) +
  geom_hline(
    yintercept = sum(provedit_comparison$value[provedit_comparison$algorithm == "Exact"]),
    linetype = "dashed"
  ) +
  geom_text(
    label = "Perfect performance", x = 2.5,
    y = sum(provedit_comparison$value[provedit_comparison$algorithm == "Exact"]) - 30,
    check_overlap = TRUE
  ) +
  scale_fill_grey(start = 0.2, end = 0.8, guide = guide_legend(reverse = T)) +
  labs(
    x = "Matching algorithm", y = "Number of correct\nmatches identified",
    fill = "True sample type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)))
```


```{r}
tibble(
  algorithm = c("Exact", "Similarity", "Static threshold", "bistro"),
  n_incorrect = c(
    nrow(provedit_exact_matches_incorrect),
    nrow(provedit_similarity_matches_incorrect),
    nrow(provedit_thresh10_matches_incorrect),
    nrow(provedit_bistro_matches_incorrect)
  )
)
```

```{r}
(provedit_bistro_compare <- bind_rows(
  provedit_exact_matches_correct_none %>%
    mutate(
      algorithm = "Exact",
      match_type = "Correct"
    ),
  provedit_similarity_matches_correct_none %>%
    mutate(
      algorithm = "Similarity",
      match_type = "Correct"
    ),
  provedit_thresh10_matches_correct_none %>%
    mutate(
      algorithm = "Static threshold",
      match_type = "Correct"
    ),
  provedit_bistro_matches_correct_none %>%
    mutate(
      algorithm = "bistro",
      match_type = "Correct"
    ),
  provedit_bistro_matches_incorrect %>%
    mutate(
      algorithm = "bistro",
      match_type = "Incorrect"
    ) %>%
    left_join(provedit_samp_composition %>% select(bloodmeal_id, multisource))
) %>%
  filter(match == "Yes") %>%
  select(-c(log10_lr, notes, thresh_low, est_noc, match)) %>%
  group_by(bloodmeal_id, human_id, multisource, incomplete, match_type) %>%
  summarize(algorithm_match = list(unique(algorithm))) %>%
  mutate(
    multisource = factor(multisource, levels = rev(unique(multisource))),
    incomplete = gsub(" profile", "", incomplete),
    type = factor(paste0(multisource, ",\n", tolower(match_type)),
      levels = c("Single-source,\ncorrect", "Multi-source,\ncorrect", "Multi-source,\nincorrect", "Single-source,\nincorrect")
    )
  ) %>%
  ggplot(aes(x = algorithm_match, fill = match_type)) +
  geom_bar() +
  scale_x_upset() +
  scale_fill_grey(start = 0.8, end = 0.2) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(x = "Algorithms that\nidentified the match", y = "Number of matches", fill = "Match type") +
  theme(
    axis.title.y = element_text(vjust = -15),
    strip.text.y = element_text(angle = 0)
  ))
```

```{r}
provedit_fig <- ggarrange(
  ggarrange(provedit_match_plot,
    provedit_mixes,
    incomplete_sens_spec,
    nrow = 1, labels = "AUTO", widths = c(1.2, 1, 1.5)
  ),
  ggarrange(provedit_compare_matches,
    provedit_bistro_compare,
    labels = c("D", "E"), nrow = 1, widths = c(1, 1)
  ),
  ncol = 1, heights = c(1, 1.1)
)
ggsave(plot = provedit_fig, filename = "figures/provedit.png", width = 10, height = 7)
```

```{r}
provedit_match_summary <- bind_rows(
  provedit_exact_matches_correct_none %>%
    group_by(bloodmeal_id, noc, multisource) %>%
    summarize(n_match = sum(match == "Yes")) %>%
    left_join(provedit_exact_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Exact"),
  provedit_similarity_matches_correct_none %>%
    group_by(bloodmeal_id, noc, multisource) %>%
    summarize(n_match = sum(match == "Yes")) %>%
    left_join(provedit_similarity_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Similarity"),
  provedit_thresh10_matches_correct_none %>%
    group_by(bloodmeal_id, noc, multisource) %>%
    summarize(n_match = sum(match == "Yes")) %>%
    left_join(provedit_thresh10_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "Static threshold"),
  provedit_bistro_matches_correct_none %>%
    group_by(bloodmeal_id, noc, multisource) %>%
    summarize(n_match = sum(match == "Yes")) %>%
    left_join(provedit_bistro_matches_incorrect %>%
      group_by(bloodmeal_id) %>%
      summarize(n_incorrect = n())) %>%
    mutate(algorithm = "bistro")
) %>%
  mutate(n_incorrect = ifelse(is.na(n_incorrect), 0, n_incorrect)) %>%
  mutate(n_missing = noc - n_match - n_incorrect) %>%
  group_by(algorithm, multisource) %>%
  summarize(
    n_samp = n(),
    n_samp_match = sum(n_match > 0),
    n_source = sum(n_match) + sum(n_incorrect) + sum(n_missing),
    n_match = sum(n_match),
    n_incorrect = sum(n_incorrect)
  )

provedit_match_summary %>%
  mutate(multisource = "All") %>%
  group_by(algorithm, multisource) %>%
  summarize(
    n_samp = sum(n_samp),
    n_samp_match = sum(n_samp_match),
    n_source = sum(n_source),
    n_match = sum(n_match),
    n_incorrect = sum(n_incorrect)
  ) %>%
  bind_rows(provedit_match_summary) %>%
  mutate(
    n_samp_match = paste0(n_samp_match, "/", n_samp, " (", signif(n_samp_match / n_samp * 100, 2), "%)"),
    n_match = paste0(n_match, "/", n_source, " (", signif(n_match / n_source * 100, 2), "%)"),
    multisource = factor(multisource, levels = c("Single-source", "Multi-source", "All")),
    algorithm = factor(algorithm, levels = rev(c("Exact", "Similarity", "Static threshold", "bistro")))
  ) %>%
  select(multisource, algorithm, n_samp_match, n_match, n_incorrect) %>%
  arrange(multisource, algorithm) %>%
  pivot_wider(
    names_from = multisource,
    values_from = c(n_samp_match, n_match, n_incorrect), names_vary = "slowest"
  ) %>%
  select(-c(`n_match_Single-source`, `n_incorrect_Single-source`, `n_incorrect_Multi-source`))
```


## Figure 3: Mosquito bloodmeals

Number of human sources

```{r}
bm_n_hu
```

Maximum similarity value between people in the database

```{r}
bm_max_similarity
```

Total number of bloodmeal samples tested

```{r}
n_distinct(bm_bistro_matches_all$bloodmeal_id)
```

Number of samples with only AMEL gene returned (not used for LR matching)

```{r}
(n_amel_only <- n_distinct(bm_exact_matches$bloodmeal_id) - n_distinct(bm_bistro_matches$bloodmeal_id))
```


Number and proportion that did not return any peaks above a 200 RFU threshold, and number remaining

```{r}
sum(bm_bistro_matches_all$notes == "all log10LRs NA or Inf;no peaks above threshold") - n_amel_only
```

Total not used for matching

```{r}
sum(bm_bistro_matches_all$notes == "all log10LRs NA or Inf;no peaks above threshold")
```

Total used for matching, and proportion

```{r}
n_distinct(bm_bistro_matches$bloodmeal_id)
n_distinct(bm_bistro_matches$bloodmeal_id) / n_distinct(bm_bistro_matches_all$bloodmeal_id)
```

Split by multisource

```{r}
bm_bistro_matches %>%
  select(bloodmeal_id, est_noc, multisource) %>%
  distinct() %>%
  mutate(tot = n()) %>%
  group_by(multisource, tot) %>%
  summarize(
    n = n(),
    prop = n / unique(tot),
    n_src = sum(est_noc)
  )
```

Estimated NOC distribution

```{r}
bm_bistro_matches %>%
  filter(multisource == "Multi-source") %>%
  select(bloodmeal_id, est_noc) %>%
  distinct() %>%
  pull(est_noc) %>%
  table()
```


```{r}
(match_noc_inc_fig <- bm_bistro_matches %>%
  group_by(bloodmeal_id, incomplete, est_noc) %>%
  mutate(
    n_matches = sum(!is.na(human_id)),
    multisource = factor(ifelse(est_noc > 1, "Multi-source", "Single-source"), levels = c("Single-source", "Multi-source")),
    incomplete = gsub(" profile", "", incomplete)
  ) %>%
  group_by(bloodmeal_id, incomplete, multisource, n_matches) %>%
  summarize() %>%
  ggplot(aes(x = incomplete, fill = as.character(n_matches))) +
  facet_grid(~multisource, scales = "free", space = "free") +
  geom_bar() +
  scale_fill_grey(start = 0.8, end = 0.2) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(
    x = "STR profile type", y = "Number of bloodmeals",
    fill = "Number of\nbistro matches\nto someone in\nthe database"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)))
```

```{r}
bm_bistro_matches %>%
  group_by(bloodmeal_id, incomplete, est_noc) %>%
  summarize(n_matches = sum(!is.na(human_id))) %>%
  group_by(incomplete) %>%
  summarize(
    n_bms = n(),
    n_geq1match = sum(n_matches > 0)
  ) %>%
  mutate(prop = n_geq1match / n_bms)

bm_bistro_matches %>%
  group_by(bloodmeal_id, incomplete, est_noc) %>%
  summarize(n_matches = sum(!is.na(human_id))) %>%
  ungroup() %>%
  summarize(
    n_tot = n(),
    n_multi = sum(est_noc > 1),
    n_possible_matches = sum(est_noc),
    n_eq1match = sum(n_matches == 1),
    n_eq2match = sum(n_matches == 2),
    n_eq3match = sum(n_matches == 3)
  ) %>%
  mutate(prop = n_eq2match / n_multi)

bm_bistro_matches %>%
  group_by(bloodmeal_id, est_noc) %>%
  summarize(n_match = sum(!is.na(human_id))) %>%
  ungroup() %>%
  summarize(
    n_src = sum(est_noc),
    n_match = sum(n_match),
    prop = n_match / n_src
  )
```

```{r}
tibble(
  algorithm = c("Exact/similarity", "Static threshold"),
  n_matches = c(
    sum(bm_exact_matches$match == "Yes"),
    sum(bm_thresh10_matches$match == "Yes")
  ),
  bistro = sum(bm_bistro_matches$match == "Yes")
) %>%
  mutate(
    diff = bistro - n_matches,
    times_increase = 1 + diff / n_matches
  )
```


```{r}
bm_exact_matches %>%
  group_by(bloodmeal_id) %>%
  summarize(n_matches = sum(!is.na(human_id))) %>%
  ungroup() %>%
  summarize(
    n_bms = n(),
    n_geq1match = sum(n_matches > 0)
  ) %>%
  mutate(prop = n_geq1match / n_bms)

bm_thresh10_matches %>%
  group_by(bloodmeal_id) %>%
  summarize(n_matches = sum(!is.na(human_id))) %>%
  ungroup() %>%
  summarize(
    n_bms = n(),
    n_geq1match = sum(n_matches > 0)
  ) %>%
  mutate(prop = n_geq1match / n_bms)
```

```{r}
bm_all_possible_long <- bm_bistro_matches %>%
  group_by(bloodmeal_id, est_noc, multisource, incomplete) %>%
  summarize() %>%
  uncount(est_noc) %>%
  group_by(bloodmeal_id) %>%
  mutate(rn = row_number()) %>%
  ungroup()

bm_exact_long <- bm_all_possible_long %>%
  left_join(bm_exact_matches %>%
    group_by(bloodmeal_id) %>%
    mutate(rn = row_number()) %>%
    ungroup()) %>%
  mutate(
    match = ifelse(is.na(match), "No", match),
    algorithm = "Exact/\nsimilarity"
  ) %>%
  select(bloodmeal_id, multisource, incomplete, match, human_id, algorithm)

bm_thresh_long <- bm_all_possible_long %>%
  left_join(bm_thresh10_matches %>%
    group_by(bloodmeal_id) %>%
    mutate(rn = row_number()) %>%
    ungroup()) %>%
  mutate(
    match = ifelse(is.na(match), "No", match),
    algorithm = "Static\nthreshold"
  ) %>%
  select(bloodmeal_id, multisource, incomplete, match, human_id, algorithm)

bm_bistro_long <- bm_all_possible_long %>%
  left_join(bm_bistro_matches %>%
    rename(
      bloodmeal_id = bloodmeal_id,
      human_id = human_id
    ) %>%
    group_by(bloodmeal_id) %>%
    mutate(rn = row_number()) %>%
    ungroup()) %>%
  mutate(
    match = ifelse(is.na(match), "No", match),
    algorithm = "bistro"
  ) %>%
  select(bloodmeal_id, multisource, incomplete, match, human_id, algorithm)

bm_comparison <- bind_rows(bm_exact_long, bm_thresh_long, bm_bistro_long)
```

```{r}
(bm_comparison_plot <- bm_comparison %>%
  mutate(
    incomplete = gsub(" ", "\n", incomplete),
    multisource = factor(multisource, levels = c("Single-source", "Multi-source")),
    algorithm = factor(algorithm, levels = c("bistro", "Static\nthreshold", "Exact/\nsimilarity"))
  ) %>%
  ggplot(aes(x = algorithm, fill = match)) +
  facet_grid(incomplete ~ multisource, scales = "free", space = "free") +
  geom_bar() +
  scale_fill_manual(values = c("grey", "grey20")) +
  scale_y_continuous(breaks = seq(0, 600, by = 50)) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(
    x = "Algorithm", y = "Number of possible matches",
    fill = "Match to\nsomeone in\nthe database"
  ) +
  theme(
    strip.text.y = element_text(angle = 0),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ))
```

```{r}
bm_fig <- ggarrange(match_noc_inc_fig + theme(legend.position = "bottom"),
  bm_comparison_plot + theme(legend.position = "bottom"),
  nrow = 1, widths = c(1, 1.5), labels = "AUTO"
)
ggsave(
  plot = bm_fig, filename = "figures/bloodmeals.png",
  width = 7, height = 6
)
```


```{r}
bm_match_summary <- bm_comparison %>%
  group_by(bloodmeal_id, multisource, algorithm) %>%
  summarize(
    n_source = n(),
    n_match = sum(match == "Yes")
  ) %>%
  group_by(algorithm, multisource) %>%
  summarize(
    n_samp = n(),
    n_samp_match = sum(n_match > 0),
    n_source = sum(n_source),
    n_match = sum(n_match)
  )

bm_match_summary %>%
  mutate(multisource = "All") %>%
  group_by(algorithm, multisource) %>%
  summarize(
    n_samp = sum(n_samp),
    n_samp_match = sum(n_samp_match),
    n_source = sum(n_source),
    n_match = sum(n_match)
  ) %>%
  bind_rows(bm_match_summary) %>%
  mutate(
    n_samp_match = paste0(n_samp_match, "/", n_samp, " (", signif(n_samp_match / n_samp * 100, 2), "%)"),
    n_match = paste0(n_match, "/", n_source, " (", signif(n_match / n_source * 100, 2), "%)"),
    multisource = factor(multisource, levels = c("Single-source", "Multi-source", "All")),
    algorithm = factor(algorithm, levels = rev(c("Exact/\nsimilarity", "Static\nthreshold", "bistro")))
  ) %>%
  select(multisource, algorithm, n_samp_match, n_match) %>%
  arrange(multisource, algorithm) %>%
  pivot_wider(
    names_from = multisource,
    values_from = c(n_samp_match, n_match), names_vary = "slowest"
  ) %>%
  select(-c(`n_match_Single-source`))
```



